A Decision Procedure for (Co)datatypes in SMT Solvers

### 2015-02-16 tex/body.tex
removed comments to self

For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we can
recursively construct its \emph{normal form} $\nf{t}$
by replacing all of the self-similar subterms of $t$ with $x$
%and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
and by normalizing the subterms.
For example, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.

@@

We now define the class of interpretations for $\thD$.
$\M( \tau )$ denotes the interpretation type $\tau$ in $\M$,
that is, a set of domain elements for that type.
For a term $t\typ{\tau}$, $\M( t ) \in \M( \tau )$ denotes the interpretation of $t$ in $\M$.
To simplify the presentation, all types are interpreted as sets of $\mu$-terms; but
only values of types in $\Codata$ may contain cycles.
%Since all values of types in $\Data$ and $\Nondata$ are acyclic,
%a typical representation for these values can be obtained by
%dropping each of its $\mu$-bindings.
%For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
%is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.

\begin{definition}[Normal Model]
\afterDot%
\label{def:norm-model}%
\rm
A model $\M$ is \emph{normal} if these conditions are met:
\begin{enumerate}
\item
%%% TODO: What does maximal set mean? Does it make sense for \Nondata?
For each type $\tau$,
$\M( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are
unique up to $\alpha$-equivalence and acyclic when $\tau \not\in \Codata$.
\item
For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
$\M( \const{C}( \vec t \,) )$ is the value
in $\M( \tau )$ that is $\alpha$-equivalent to
$\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$, where $x$ is fresh.
%%% TODO: I'm a bit worried to see a semantic value \M under a syntactic
%%% constructor. I know our models consist of \mu-terms, but is this intended?
\item
For all selector terms $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
$\M\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
in $\M( \tau )$ that is $\alpha$-equivalent to
$\interp{\M( t )}{u_k}$,
assuming $\M( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$.
\end{enumerate}
\end{definition}

>>>>

For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we can
construct its \emph{normal form} $\nf{t}$
by replacing all of the self-similar subterms of $t$ with $x$
%and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
and by recursively normalizing $t$'s subterms.
For example, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.

@@

We now define the class of interpretations for $\thD$.
$\M( \tau )$ denotes the interpretation type $\tau$ in $\M$---that is, a set of
domain elements for that type.
For a term $t\typ{\tau}$, $\M( t ) \in \M( \tau )$ denotes the interpretation of $t$ in $\M$.
%To simplify the presentation,
All types are interpreted as sets of $\mu$-terms, but
only values of types in $\Codata$ may contain cycles.
%Since all values of types in $\Data$ and $\Nondata$ are acyclic,
%a typical representation for these values can be obtained by
%dropping each of its $\mu$-bindings.
%For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
%is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.

\begin{definition}[Normal Model]
\afterDot%
\label{def:norm-model}%
\rm
A model $\M$ is \emph{normal} if these conditions are met:
\begin{enumerate}
\item
For each type $\tau$,
$\M( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are
unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
\item
For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
$\M( \const{C}( \vec t \,) )$ is the value
in $\M( \tau )$ that is $\alpha$-equivalent to
$\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$, where $x$ is fresh.
\item
For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
$\M\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
in $\M( \tau )$ that is $\alpha$-equivalent to
$\interp{\M( t )}{u_k}$,
assuming $\M( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$.
\end{enumerate}
\end{definition}
